---
title: "Inspection of causalities in voting preferences"
author: "Anastasiia Livochka, Yarka Lyba"
date: "27 Apr 2019"
output:
  html_document:
    number_sections: yes
    theme: readable
    toc: yes
---

# Loadind packages
```{r}
library(gridExtra)
library(base)
library(factoextra)
library(corrplot)
library(dplyr)
library(randomForest)
library(ggplot2)
```

# Loadind data
```{r}
y2 <- read.csv(file="data2019_prepared.csv")
y2$X <- NULL
```

# Visualization of voting outcomes in dataset
Voting outcomes in or data coincide with real outcome of elections 2019 in Ukraine.
```{r}
pie <- ggplot(y2, aes(x = "", fill = factor(zelensk))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="Zelenskiy?")
pie + coord_polar(theta = "y", start=0)
```
# Random Forest model and feature selection
Data consist of 96 variables, so we need to select the most inportant features for solve the problem of overfitting. 

```{r}
electionsRF <- randomForest(factor(zelensk)~.-zelensk, y2,ntree=150)
electionsRF
```
```{r}
importance    <- importance(electionsRF)

varImportance <- data.frame(Variables = row.names(importance), 
                            Importance = round(importance[ ,'MeanDecreaseGini'],2)) 
top20 <- dplyr::arrange(varImportance, desc(Importance))[1:20, ]
top20$Variables
rankImportance <- top20 %>%
  mutate(Rank = paste0('#',dense_rank(desc(Importance))))
ggplot(rankImportance, aes(x = reorder(Variables, Importance), 
                           y = Importance, fill = Importance)) +
  geom_bar(stat='identity') + 
  geom_text(aes(x = Variables, y = 0.5, label = Rank),
            hjust=0, vjust=0.55, size = 4, colour = 'red') +
  labs(x = 'Variables') +
  coord_flip() + 
  theme_classic()
```
Selecting the most important variables for inspecting the causalities
```{r}
y2.important <- dplyr::select(y2, zelensk, trust_president, west, gas_cheap, east, trust_parl, trust_army, abroad_west, trust_church, trust_police, r_u_same, rus_language, alco, gov_all, east_komprom, greek_cath, move_forever, trust_family,   trust_edu, trust_banks, life_satisf)
```

Running the logit model with targeting whether the person vote for Zelenskiy
```{r}
election_model <- glm(factor(zelensk) ~. -zelensk, y2.important, family = binomial(logit))
summary(election_model)
```
# Clustering people by their political views
We decided to try clusterizing Ukrainian electorate by theit opinion about following things:
```{r}
y2.political <- dplyr::select(y2, gas_cheap, r_u_same, gov_all,  zelensk, trust_ZMI, trust_army, trust_court, trust_church,  trust_president, trust_parl, abroad_east, abroad_both, abroad_west, east_komprom, east_frozen, east_aggres, travel_ukr)
```
```{r}
fviz_nbclust(y2.political, kmeans,method = "gap_stat")
```

We got optimal number of clusters equal to 4 via gap statistics method. You can find more here: https://www.datanovia.com/en/lessons/determining-the-optimal-number-of-clusters-3-must-know-methods/. 
```{r}
political_clusters <- kmeans(y2.political, 4)
political_clusters$size
```
You can see that the number of people in clusters are pretty well balanced.


## Visualization of clusters
Firstly let's look on voting outcomes in clusters:
```{r}
 zel_1 <- y2.political[political_clusters$cluster==1,] %>%
  ggplot(aes(x = "", fill = factor(zelensk))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0) 

zel_2 <- y2.political[political_clusters$cluster==2,] %>%
  ggplot(aes(x = "", fill = factor(zelensk))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0)

zel_3 <- y2.political[political_clusters$cluster==3,] %>%
  ggplot(aes(x = "", fill = factor(zelensk))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0) 
zel_4 <- y2.political[political_clusters$cluster==4,] %>%
  ggplot(aes(x = "", fill = factor(zelensk))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0)
grid.arrange(zel_1, zel_2, zel_3, zel_4, nrow=2, top="Zelenskiy?")
```
<br/>**Fact1: The biggest fraction of people who support Zelenskiy is in the first cluster, the fourth cluster consist mainly of Poroshenko supporters**
```{r}
 r_u1 <- y2.political[political_clusters$cluster==1,] %>%
  ggplot(aes(x = "", fill = factor(r_u_same))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0) 

r_u2 <- y2.political[political_clusters$cluster==2,] %>%
  ggplot(aes(x = "", fill = factor(r_u_same))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0) 

 r_u3 <- y2.political[political_clusters$cluster==3,] %>%
  ggplot(aes(x = "", fill = factor(r_u_same))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0) 
 r_u4 <- y2.political[political_clusters$cluster==4,] %>%
  ggplot(aes(x = "", fill = factor(r_u_same))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0) 
grid.arrange(r_u1,  r_u2, r_u3, r_u4, nrow=2, top="Does Russians and Ukrainians form one nation?")
```
<br/>**Fact2: The biggest fraction of people who consiber Ukrainian and Russian as the same public is in the first cluster, the smallest in the fourth**
```{r}
cheap1 <- y2.political[political_clusters$cluster==1,] %>%
  ggplot(aes(x = "", fill = factor(gas_cheap))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0) 

cheap2 <- y2.political[political_clusters$cluster==2,] %>%
  ggplot(aes(x = "", fill = factor(gas_cheap))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0) 

 cheap3 <- y2.political[political_clusters$cluster==3,] %>%
  ggplot(aes(x = "", fill = factor(gas_cheap))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0) 
 cheap4 <- y2.political[political_clusters$cluster==4,] %>%
  ggplot(aes(x = "", fill = factor(gas_cheap))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0) 
grid.arrange(cheap1,  cheap2, cheap3, cheap4, nrow=2, top="Some cheap gas?")
```
<br/>**Fact3:  The biggest fraction of people who support artificial reduction of tarrifs is in the first cluster, the smallest is in the fourth**
```{r}
all1 <- y2.political[political_clusters$cluster==1,] %>%
  ggplot(aes(x = "", fill = factor(gov_all))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0) 

all2 <- y2.political[political_clusters$cluster==2,] %>%
  ggplot(aes(x = "", fill = factor(gov_all))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0) 

 all3 <- y2.political[political_clusters$cluster==3,] %>%
  ggplot(aes(x = "", fill = factor(gov_all))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0) 
 all4 <- y2.political[political_clusters$cluster==4,] %>%
  ggplot(aes(x = "", fill = factor(gov_all))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0) 
grid.arrange(all1,  all2, all3, all4, nrow=2, top="All desires of the public?")
```
<br/>**Fact3:  The biggest fraction of people who support the policy of complying with all desires of the public is in the first public, the smallest is in the fourth**
```{r}
eu1 <- y2.political[political_clusters$cluster==1,] %>%
  ggplot(aes(x = "", fill = factor(abroad_west))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0) 

eu2 <- y2.political[political_clusters$cluster==2,] %>%
  ggplot(aes(x = "", fill = factor(abroad_west))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0) 

 eu3 <- y2.political[political_clusters$cluster==3,] %>%
  ggplot(aes(x = "", fill = factor(abroad_west))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0) 
 eu4 <- y2.political[political_clusters$cluster==4,] %>%
  ggplot(aes(x = "", fill = factor(abroad_west))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="") + coord_polar(theta = "y", start=0) 
grid.arrange(eu1,  eu2, eu3, eu4, nrow=2, top="Travelling to the Europe?")
```
<br/>**Fact4:  The biggest fraction of people who travel to the Europe is in the fourth cluster, the smallest is in the first one** <br/>
More conclusions and interpretation you can find in the report attached at https://github.com/livochka/youth-opinion-Ukraine
